Challenges to transboundary fisheries management in North America under climate change

### 2020-11-26 Ecol/Manuscript_Draft.Rmd
changed reference for chapters

In 1982, the United Nations Law of the Sea Convention (UNCLOS) formalized the concept of exclusive economic zones (EEZs) creating what we know today as shared stocks [@UnitedNations:1986tl], i.e. stocks that migrate between countries EEZs (known as transboundary stocks) or between EEZs and the high seas (also called straddling stocks) [@Song:2017iua]. Today, an estimated 347 [@Teh:2015gd] to 1500 [@Caddy:1997ue] fish stocks cross national borders, some of them jointly managed by two or more countries. These stocks are responsible for almost 50% of these countries total fish catches [@Teh:2015gd]. Under Article 63, UNCLOS incentives actions to cooperate on the management of shared stocks [@UnitedNations:1986tl] as often management success depends on effective cooperation between parties [@Munro:2002uf; @Sumaila:2013hv]. Since the definition of shared stocks, game theory has been one of the most common approaches used to analyze the management of these type of stocks. However, shared stocks' management can be convoluted due to the participation of several fishing “players”, different countries and sometimes jurisdictions within a country, the migration patterns of the stock, and their abundance fluctuation within space and time [@Munro:2002uf, Engler and Saunders *in press*]. In addition, international treaties might not be prepared to address the effects that climate change will bring to shared fish stocks (Engler and Saunders *in press*, Koubrak and VanderZwaag, *in press*).

@@

The ocean is getting warmer [@Rheim:2013fv], less oxygenated [@Schmidtko:2017cl], and increasing in acidity [@Ross:2011kd; @IPCC:2019tn]. To cope with these changes in ocean biophysic properties, marine species, including shared fish stocks, have been shifting their distribution towards the poles and/or deeper waters [@Poloczanska:2016kk]. As climate change reshapes the ocean's environment worldwide [@Gattuso:2015jz], shared fisheries' delicate governance is threatened as new migration patterns may arise [@Miller:2013iv; @Pinsky:2018cb], historic distribution and abundances might shift [@Cheung:2010dt], and species basic natural traits may modify [@Pauly:2017hpa]. Catches of shared stocks like tunas, have significantly increased in some regions such as the subtropical Atlantic and western Pacific Oceans and are projected to continue [@MonllorHurtado:2017cm; @ErauskinExtramiana:2019fw]. Multiple shared species in North America have been observed to shift in distribution following changes in optimal conditions such as sockeye salmon (*Oncorhynchus nerka*) [@McDaniels:2010vn], Atlantic cod [@Pershing:2015gq], and flounders [@Pinsky:2012kq]. Moreover, these shifts are projected to continue towards the end of the $21^{st}$ century [@Cheung:2018hz]. As a result, some countries or management jurisdictions may see more shared fisheries and their catches shifting into their waters while others will stand to lose [@Oremus:2020id; @Pinsky:2018cb]. Nevertheless, management rules for shared stocks (e.g., quota or spatial delimitation) are often determined based on current and/or historic knowledge of the stock's distribution and do not consider future shifts in distributions [@FredstonHermann:2018kp].

>>>>

In 1982, the United Nations Law of the Sea Convention (UNCLOS) formalized the concept of exclusive economic zones (EEZs) creating what we know today as shared stocks [@UnitedNations:1986tl], i.e. stocks that migrate between countries EEZs (known as transboundary stocks) or between EEZs and the high seas (also called straddling stocks) [@Song:2017iua]. Today, an estimated of 347 [@Teh:2015gd] to 633 (Chapter 2) fish stocks cross national borders, some of them jointly managed by two or more countries. These stocks are responsible for more than 70% of these countries total fish catches (Chapter 2). Under Article 63, UNCLOS incentives actions to cooperate on the management of shared stocks [@UnitedNations:1986tl] as often management success depends on effective cooperation between parties [@Munro:2002uf; @Sumaila:2013hv]. Since the definition of shared stocks, game theory has been one of the most common approaches used to analyze the management of these type of stocks. However, shared stocks' management can be convoluted due to the participation of several fishing “players”, different countries and sometimes jurisdictions within a country, the migration patterns of the stock, and their abundance fluctuation within space and time [@Munro:2002uf, Engler and Saunders *in press*]. In addition, international treaties might not be prepared to address the effects that climate change will bring to shared fish stocks (Engler and Saunders *in press*, Koubrak and VanderZwaag, *in press*).

@@

The ocean is getting warmer [@Rheim:2013fv], less oxygenated [@Schmidtko:2017cl], and increasing in acidity [@Ross:2011kd; @IPCC:2019tn]. To cope with these changes in ocean biophysic properties, marine species, including shared fish stocks, have been shifting their distribution towards the poles and/or deeper waters [@Poloczanska:2016kk]. As climate change reshapes the ocean's environment worldwide [@Gattuso:2015jz], shared fisheries' delicate governance is threatened as new migration patterns may arise [@Miller:2013iv; @Pinsky:2018cb], historic distribution and abundances might shift [@Cheung:2010dt], and species basic natural traits may modify [@Pauly:2017hpa]. Catches of shared stocks like tunas, have significantly increased in some regions such as the subtropical Atlantic and western Pacific Oceans and are projected to continue [@MonllorHurtado:2017cm; @ErauskinExtramiana:2019fw]. Multiple shared species in North America have been observed to shift in distribution following changes in optimal conditions such as sockeye salmon (*Oncorhynchus nerka*) [@McDaniels:2010vn], Atlantic cod [@Pershing:2015gq], and flounders [@Pinsky:2012kq]. Moreover, these shifts are projected to continue towards the end of the $21^{st}$ century [@Cheung:2018hz]. As a result, some countries or management jurisdictions may see more shared fisheries and their catches shifting into their waters while others will stand to lose [@Oremus:2020id; @Pinsky:2018cb, Chapter 3]. Nevertheless, management rules for shared stocks (e.g., quota or spatial delimitation) are often determined based on current and/or historic knowledge of the stock's distribution and do not consider future shifts in distributions [@FredstonHermann:2018kp].
